### 2023-12-03 header.tex
Added Jay Sanjay as coauthor

% **************GENERATED FILE, DO NOT EDIT**************

\title{Developing Observational Health Research Capacity in the Julia Ecosystem}

\author[1]{Jacob S. Zelko}
\author[3]{Fareeda Abdelazeez}
\author[1, 2]{Malina Hy}
\author[1, 2]{Varshini Chinta}
\author[3]{Jay Sanjay}
\affil[1]{Georgia Tech Research Institute}
\affil[2]{Georgia Institute of Technology}
\affil[3]{Foobar}

\keywords{Observational Health, OMOP Common Data Model, Database Management, Characterization}

\hypersetup{
pdftitle = {Developing Observational Health Research Capacity in the Julia Ecosystem},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Jacob S. Zelko, Fareeda Abdelazeez, Malina Hy, Varshini Chinta, Jay Sanjay},
pdfkeywords = {Observational Health, OMOP Common Data Model, Database Management, Characterization},
}