### 2023-12-31 header.tex
Renamed paper and added author information

% **************GENERATED FILE, DO NOT EDIT**************

\title{The JuliaHealth Observational Health Research Subecosystem}

\author[5]{Jacob S. Zelko}
\author[1, 2]{Malina Hy}
\author[1, 2]{Varshini Chinta}
\author[3]{Fareeda Abdelazeez}
\author[4]{Jay Sanjay}
\affil[1]{Georgia Tech Research Institute}
\affil[2]{Georgia Institute of Technology}
\affil[3]{Zewail University}
\affil[4]{Foobar}
\affil[5]{Northeastern University}

\keywords{Observational Health, OMOP Common Data Model, Database Management, Characterization}

\hypersetup{
pdftitle = {The JuliaHealth Observational Health Research Subecosystem},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Jacob S. Zelko, Malina Hy, Varshini Chinta, Fareeda Abdelazeez, Jay Sanjay},
pdfkeywords = {Observational Health, OMOP Common Data Model, Database Management, Characterization},
}